\LARGE \bf Mean Estimation from Adaptive One-bit Measurements

### 2017-10-09 Writeups/onebit_adaptive_Allerton_V3.tex
small changes to final version

Estimation performance of information from data collected and processed by multiple units may be limited due to communication constraints between these units. 
%For example, emerging sensor network technologies in medicine and ``smart cities" use many low-cost sensors to collect large amounts of data and transmit it to remote locations. These sensors must operate under severe power restrictions, hence they are limited by their communication bandwidth. 
For example, consider large-scale sensor arrays where information is collected through multiple sensors and transmitted to a central estimation unit. In this scenario, the ability to estimate a particular parameter from the data is dictated not only by the quality of observations and their number, but also by available rate of communication between the sensors and the central estimator. The question that we ask is to what extent a parametric estimation task is affected by communication constraints, and what are the fundamental performance limits in estimating a parameter subject to these restrictions. In this paper we answer this question in a particular setting: the estimation of the mean $\theta$ of a normal distribution with variance $\sigma^2$ under the constraint that only a single bit can be communicated on each sample $X_n$ from this distribution. As it turns out, the ability to share information among different samples before committing on each single-bit message dramatically affects the performance in estimating $\theta$. We therefore distinguish in among three settings:
 \begin{itemize}
 \item[(i)] \emph{Centralized} encoding: all $n$ encoders confer and produce a single $n$ bit message which is a function of $X_1,\ldots,X_n$. 
 \item[(ii)] \emph{Adaptive} or \emph{sequential} encoding: the $n$th encoder observes $X_n$ and the $n-1$ previous single bit messages.
 \item[(iii)] \emph{Distributed} encoding: the output of the $n$th encoder is a single bit that is only a function of $X_n$.
 \end{itemize}
Clearly, as far as information sharing is concerned, settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). We measure the estimation performance by the mean squared error (MSE) risk. We are interested in particular in the \emph{asymptotic relative efficiency} (ARE) of estimators in the constrained setting relative to the MSE attained by the sample mean, which is the minimax estimator in estimating without one-bit constraint and its MSE decreases as $\sigma^2/n+O(n^{-2})$. \\

In setting (i) the estimator can evaluate the sample mean and then communicate it using $n$ bits. This strategy leads to MSE behavior of $\sigma^2/n + O(2^{-n})$. Therefore, the ARE in this setting is $1$. Namely, asymptotically, there is no loss in performance due to the communication constraint under centralized encoding. In this work we show that a similar result does not hold even in setting (ii): the ARE of any adaptive estimation scheme is at least $\pi/2$. Namely, the single-bit per sample constraint incurs a minimal penalty of at least $1.57$ in the number of samples compared to an unconstrained estimator or to the optimal estimator in setting (i). In addition to this negative statement, we provide an estimator that attains this ARE. In other words, we show that the lower bound of $\pi/2$ on the ARE is tight, and that it is attained regardless of the particular realization of $\theta$ or the radius of the parameter space from which it is taken. Clearly, the minimal penalty on the efficiency of $\pi/2$ also holds under setting (iii), although the question whether this efficiency is achievable (or otherwise, what is the minimal ARE) remains open. 
\\

@@

We also note that our settings (ii) and (iii) are special cases of 
\cite{zhang2013information} that consider adaptive and distributed estimation protocols for $m$ machines, each has access to $n/m$ independent samples. The main result of \cite{zhang2013information} are bounds on the estimation error as a function of the number of bits $R$ each machine uses for communication. The specialization of their result to our setting by, taking $m=n$ and $R=1$, leads to looser bounds then $\sigma^2\pi/(2n)$ for cases (ii) and (iii). 
Other related works include statistical inference under multiterminal data compression \cite{han1987hypothesis, zhang1988estimation}, although their results apply only to setting (i). The counterpart of setting (iii) in the case of hypothesis testing was considered in \cite{52470}.
%
Estimation problem subject to the one-bit quantization constraint were also considered recently in compressed sensing \cite{boufounos20081,baraniuk2017exponential} and in MIMO detection in wireless communication \cite{singh2009limits}. \\

@@

In this work we are concerned with the MSE risk defined as
\begin{equation}
\label{eq:error_def}
\mathbb E\left(\widehat{\theta}_n - \theta \right)^2,
\end{equation}
where the expectation is taken with respect to the distribution of $X^n$ and the prior distribution $\pi(\theta)$. \\

The main problem we consider is the minimization of \eqref{eq:error_def} over all encoding and estimation strategies, and the characterization of its minimal value as a function of $n$. This minimization is the combination of the following two procedures: (1) selecting the $i$th message $M_i$ based on past messages and current observation $X_i$, and (2) estimating $\theta$ given messages $M^n$. 
%
We are interested in particular on the increase in sample complexity compared to the vanilla mean estimation without one-bit constraint. For this reason, we consider the limit
\begin{equation}
{\sigma^2} n \mathbb E\left(\widehat{\theta}_n - \theta \right)^2,
\label{eq:relative_efficiency}
\end{equation}
which is the ratio between the MSE of the sample mean and the MSE of the estimator $\widehat{\theta}_n$. Note that if the limit $\mathbb E\left(\widehat{\theta}_{n} - \theta \right)^2$ exists and finite, than \eqref{eq:relative_efficiency} is the ARE of $\widehat{\theta}_n$.

@@

Before deriving our main results, we comment on the relation between our setting and the remote multiterminal source coding problem.

\section{Relation to Remote Multiterminal Source Coding \label{sec:ceo}}
In this section we draw a connection between our general problem of mean estimation from one-bit samples and the remote multiterminal source coding problem \cite{berger1996ceo}, also known as the CEO problem. The setting of the CEO includes $n$ encoders, each has access to a noisy version of a random source sequence. 
% $\left\{ U_j,\,j=1,2,\ldots \right\}$. 
The $i$th encoder observes $k$ noisy source symbols and transmit $R_i k$ bits to a central estimator. \par
Assuming that $\theta$ is drawn once from the prior $\pi(\theta)$, our mean estimation problem from one-bit samples under distributed encoding (setting (iii) in the Introduction) corresponds to the Gaussian CEO setting with $k=1$ source realization: the $i$th encoder uses $R_i=1$ bits to transmit a message that is a function of $X_i = \theta + \sigma Z_i$, where $Z_i$ is standard normal. As a result, the MSE in the optimal source coding scheme for the CEO with $n$ terminals at rates $R_1 = \ldots = R_n = 1$, a Gaussian observation noise at each estimate of variance $\sigma^2$, and an arbitrary number of $k$ independent draws of $\theta$, provides a lower bound on the MSE distortion in estimating $\theta$ in the distributed encoding setting. Note that the difference between the CEO and ours is in the fact encoders in the CEO describe $k$ realizations of $\theta$ using $k$ bits with MSE averaged over these realization, whereas our setting only allows $k=1$. 
 \\

@@

\begin{proof}
We consider the expression \cite[Eq. 10]{chen2004upper} that provides the minimal distortion $D^\star$ in the CEO with $L$ observers and under a total sum-rate $R_\Sigma = R_1 + \ldots +R_L$:
\begin{equation} \label{eq:ceo_optimal_sumrate}
R_{\Sigma} = \frac{1}{2} \log^+ \left[ \frac{\sigma_\theta^2}{D^\star} \left( \frac{D^\star L}{ D^\star L - \sigma^2 + D^\star \sigma^2 / \sigma_\theta^2 }\right)^L  \right].
\end{equation}
Assuming $R_\Sigma = n$ and $L=n$, we get
\begin{equation} \label{eq:ceo_optimal_sumrate}
n = \frac{1}{2} \log_2 \left[ \frac{\sigma_\theta^2}{D^\star} \left(\frac{ D^\star n }{D^\star n - \sigma^2 + D^\star \sigma^2/\sigma_\theta^2 }  \right)^n  \right].
\end{equation}
The value of $D^\star$ that satisfies the equation above describes the MSE under an optimal allocation of the sum-rate $R_\Sigma = n$ among the $n$ encoders. Therefore, $D^\star$ provides a lower bound to the CEO distortion with $R_1=\ldots,R_n = 1$ and hence a lower bound to the minimal MSE in estimating $\theta$ in the distributed encoding setting. By considering $D^\star$ in \eqref{eq:ceo_optimal_sumrate} as $n\rightarrow \infty$, we see that 
\[
D^\star = \frac{ 4\sigma^2 }{3n + 4 \sigma^2 / \sigma_\theta^2 } + o(n^{-1}) =  \frac{4\sigma^2}{3n} + o(n^{-1}). 
\]
\end{proof}
Since the lower bound \eqref{eq:ceo_bound}  was derived assuming optimal allocation of the $n$ bits among the encoders, it may seem as if a tighter bound can be obtained in our case by considering the CEO distortion with $R_1=\ldots = R_n = 1$. However, an upper bound for the CEO distortion under this condition follows from \cite[Prop. 5.2]{KipnisRini2017}, and leads to
\[
D_{CEO} \leq  \left( \frac{1}{\sigma_\theta^2} +  \frac{3n}{4\sigma^2 + \sigma_\theta^2} \right)^{-1}   =
\frac{4 \sigma^2}{3n} +  \frac{\sigma_\theta^2}{3n} + O(n^{-2}),
\]
which is equivalent to \eqref{eq:ceo_bound} when $\sigma_\theta$ goes to zero. \par
We conclude therefore that the difference between the MSE lower bound \eqref{eq:ceo_bound} and the actual MSE in the distributed encoding setting, is attributed exclusively to the ability to perform coding over blocks. Namely, the freedom to consider an arbitrary number $k$ of independent realizations of the problem, versus only one in ours. In other words, it is the ability to exploit the geometry of a high-dimensional product space, rather than the distributed nature of the problem, that distinguishes between the CEO distortion and our problem of mean estimation from one-bit samples.\par
In the next section we show that the ARE in adaptive encoding setting does not exceeds $\pi/2$, and thus provides a tighter lower bound for the distributed encoding setting than $4/3$ of  \eqref{eq:ceo_bound}.

\section{Results \label{sec:sequential}}
The first main result of this paper, as described in Thm.~\ref{thm:adpative_lower_bound} below, states that the ARE of any adaptive estimator cannot be lower than $\pi/2$. Next, we provide a particular adaptive estimation scheme and show in Thm.~\ref{thm:sgd} that its efficiency is $\pi/2$. Finally, in Thm.~\ref{thm:opt_one_step}, we provide an adaptive estimation scheme that is one-step optimal in the sense that at each step $i$, the message $M_i$ that minimizes the MSE given $X_i$ and the previous $M^{i-1}$ messages is chosen. While it is not clear whether the efficiency of this last scheme is $\pi/2$, numerical simulations suggests that the MSE of this scheme times $n$ also converges to $\pi/2$ faster than the first scheme.

@@

\subsection{A lower bound on adaptive one-bit schemes}
Our first results asserts that the ARE \eqref{eq:relative_efficiency} of any adaptive estimation scheme is bounded from below by $\pi/2$, as follows from the following theorem:
\begin{thm}[minimal relative effeciency] \label{thm:adpative_lower_bound}
Let $\widehat{\theta}_n$ be any estimator of $\theta$ in the adaptive setting of Fig.~\ref{fig:sequential}. Assume that $\Theta$ is an interval and that the density of the prior $\pi(\theta)$ converges to zero at the endpoints of $\Theta$. Then
\[
\mathbb E\left[ (\theta-\theta_n)^2 \right] \geq  \frac{\pi \sigma^2 }{2n +\pi \sigma^2  I_0}  =  \frac{\pi}{2n}\sigma^2+O(n^{-2}),
\]
where 
\[
I_0 = \mathbb E \left( \frac{d}{d\theta} \log \pi (\theta) \right)^2
\]
is the Fisher information with respect to a location model in $\theta$. 
\end{thm}

@@

For the estimator defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg} we have the following results:
\begin{thm} \label{thm:sgd}
The sequence $\widehat{\theta}_n$ of \eqref{eq:sgd_est} satisfies
\begin{enumerate}
\item[(i)]
\[
\sqrt{n} \left( \widehat{\theta}_n - \theta \right) \overset{d}{\rightarrow} \mathcal N \left(0,  \pi \sigma^2 /2 \right).
\]
\item[(ii)] In addition to the conditions above, assume that $\gamma_n = o(n^{-2/3})$ and $\sum_{n=1}^\infty \gamma_n = \infty$ (e.g., $\gamma_n = n^{-\beta}$ with $2/3<\beta<1$). Then
\[
\lim_{n\rightarrow \infty} n\mathbb E \left[ \left(\theta-\widehat{\theta}_n \right)^2 \right] = \frac{\pi}{2} \sigma^2 . 
\]
\end{enumerate}

@@

Thm.~\ref{thm:sgd} implies that the estimator $\widehat{\theta}_n$, defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg}, attains the minimal ARE as established by Thm.~\ref{thm:adpative_lower_bound}.
\par
Note that $\theta_0$ is not explicitly defined 
in equation \eqref{eq:sgd_est}. While a reasonable initialization is $\theta_0 = \mathbb E [\theta]$,  Thm.~\ref{thm:sgd} implies that the asymptotic behavior of the estimator is indifferent to this initialization. Thus, the optimal efficiency is attained regardless of the prior distribution on $\theta$ or the size of the parameter space $\Theta$. Nevertheless, the bound in Thm.~\ref{thm:adpative_lower_bound} suggests that the non-asymptotic estimation error can be significantly reduced whenever the location information $I_0$ is large. In contrast, the one-step optimal scheme presented in the following subsection updates the prior distribution on $\theta$ given all information gathered until step $n-1$ to provide the step $n$ estimate and prior. In particular, the this scheme exploit the prior information on $\theta$ provided by $\pi(\theta)$.

\subsection{One-step optimal estimation}
We now consider an estimation scheme that posses the property of \emph{one-step optimality}: at each step $i$, the $i$th encoder designs the detection region $M_i^{-1}(1)$ such that the MSE given $M^i$ is minimal. In other word, this scheme designs the messages in a greedy manner, such that the MSE at step $i$ is minimal given the current state of the estimation described by $M^{i-1}$. \\

@@

\begin{proof}
The proof is completed by the following two lemmas, proofs of which can be found in the Appendix:
\begin{lem} \label{lem:unique}
Let $f(x)$ be a log-concave probability density function. Then the equation 
\begin{equation}
\label{eq:lem_fixed_point}
2x = \frac{\int_x^\infty uf(u)du}{\int_x^\infty f(u)du} + \frac{\int_{-\infty}^x uf(u)du}{\int_{-\infty}^x f(u)du} 
\end{equation}
has a unique solution.
\end{lem}
\begin{lem} \label{lem:adaptive}
Let $U$ be an absolutely continuous random variable with
pdf $P(du)$. Then the one-bit message $M^\star\in \{-1,1\}$ that minimizes
\[
\int \left( u - \mathbb E[U|M(u)]  \right)^2 P(du)
\]
is given by
\[
M^\star  =  \sgn(U - \tau),
\]
where $\tau$ is the unique solution to
 \[
2 \tau = \frac{\int_{\tau}^\infty u P(du)} {\int_{\tau}^\infty P(du)} + \frac{\int_{-\infty}^{\tau} u P(du)}{\int_{-\infty}^{\tau} P(du)}.
\]
\end{lem}
\end{proof}

Thm.~\ref{thm:opt_one_step} suggests the following adaptive encoding and estimation scheme: 
\begin{itemize}
\item Initialization: set $P_0(t) = \pi(\theta)$.
\item For $n\geq 1$:
\begin{enumerate}
\item Update the prior as
\begin{align}
P_n(t) = & P(\theta=t |M^n) \\
& = \frac{ P\left( \theta=t | M^{n-1} \right) P(M_n | \theta = t , M^{n-1})  } { P(M_n | M^{n-1} )} \nonumber \\ 
& = \alpha_n  P_{n-1}(t) \Phi\left(M_n \frac{ t - \tau_{n-1} }{\sigma} \right), \label{eq:density_update}
\end{align}
where $\alpha_n$ is a normalization coefficient that equals to
\[
\alpha_n = \left(\int_{\mathbb R} P_{n-1}(t) \Phi\left(M_n \frac{t- \tau_{n-1} }{\sigma} \right)  dt \right)^{-1}. 
\]
\item The $n$th estimate for $\theta$ is the conditional expectation of $\theta$ given $M^n$, namely
\begin{equation}
\theta_n = \mathbb E \left[ \theta| M^n\right] = \int_{-\infty}^\infty t P_n(t) dt. \label{eq:estimator_update}
\end{equation}
\item Solve equation \eqref{eq:fixed_point} with the updated prior $P_n(t)$ instead of $\pi(d\theta)$. Note that since the standard normal cdf $\Phi(x)$ is log-concave, the updated prior $P_n(t)$ remains log-concave and thus a unique solution to \eqref{eq:fixed_point} is guaranteed by Lem.~\ref{lem:unique}. 
\item Update the $(n+1)$th message as
\begin{equation}\label{eq:message_update}
M_{n+1} = \sgn(X_{n+1}-\tau_n)
\end{equation}
\end{enumerate}
\end{itemize}
Since equation \eqref{eq:fixed_point} has no analytic solution, it is hard to derive the asymptotic behavior of the estimator defined by \eqref{eq:estimator_update} and \eqref{eq:message_update}. We conjecture that it attains the asymptotic relative efficiency of $\sigma^2\pi/2$ as can be observed from the numerical simulation illustrated in Fig.~\ref{fig:adaptive_error}. Also shown in Fig.~\ref{fig:adaptive_error} are the normalized MSE of the asymptotically optimal estimator defined by \eqref{eq:sgd_alg} and \eqref{eq:sgd_est}, as well as the MSE achieved by the sample mean for the same sample realization.

@@

\subsection*{Proof of Thm.~\ref{thm:opt_one_step}}
In this subsection we prove Lemmas \ref{lem:adaptive} and \ref{lem:unique} that lead to Thm.~\ref{thm:opt_one_step}.

>>>>

The performance in estimating information from data collected and processed by multiple units may be limited due to communication constraints between these units. 
%For example, emerging sensor network technologies in medicine and ``smart cities" use many low-cost sensors to collect large amounts of data and transmit it to remote locations. These sensors must operate under severe power restrictions, hence they are limited by their communication bandwidth. 
For example, consider large-scale sensor arrays where information is collected at multiple physical locations and transmitted to a central estimation unit. In this scenario, the ability to estimate a particular parameter from the data is dictated not only by the quality of observations and their number, but also by the available rate for communication between the sensors and the central estimator. The question that we ask is to what extent a parametric estimation task is affected by these communication constraints, and what are the fundamental performance limits in estimating a parameter subject to these restrictions. In this paper we answer this question in a particular setting: the estimation of the mean $\theta$ of a normal distribution with variance $\sigma^2$ under the constraint that only a single bit can be communicated on each sample $X_n$ from this distribution. As it turns out, the ability to share information among different samples before committing on each single-bit message dramatically affects the performance in estimating $\theta$. We therefore distinguish among three settings:
 \begin{itemize}
 \item[(i)] \emph{Centralized} encoding: all $n$ encoders confer and produce a single $n$ bit message which is a function of $X_1,\ldots,X_n$. 
 \item[(ii)] \emph{Adaptive} or \emph{sequential} encoding: the $n$th encoder observes $X_n$ and the $n-1$ previous single bit messages.
 \item[(iii)] \emph{Distributed} encoding: the output of the $n$th encoder is a single bit that is only a function of $X_n$.
 \end{itemize}
Clearly, as far as information sharing is concerned, settings (iii) is a more restrictive version of (ii) which is more restrictive than (i). We measure the estimation performance by the mean squared error (MSE) risk. We are interested in particular in the \emph{asymptotic relative efficiency} (ARE) of estimators in the constrained setting relative to the MSE attained by the empirical mean of the samples, which is the minimax estimator in estimating without one-bit constraint and its MSE decreases as $\sigma^2/n+O(n^{-2})$. \\

In setting (i) the estimator can evaluate the empirical mean of the samples and then communicate it using $n$ bits. This strategy leads to MSE behavior of $\sigma^2/n + O(2^{-n})$. Therefore, the ARE in this setting is $1$. Namely, asymptotically, there is no loss in performance due to the communication constraint under centralized encoding. In this work we show that a similar result does not hold even in setting (ii): the ARE of any adaptive estimation scheme is at least $\pi/2$. Namely, the single-bit per sample constraint incurs a minimal penalty of at least $1.57$ in the number of samples compared to an unconstrained estimator or to the optimal estimator in setting (i). In addition to this negative statement, we provide an estimator that attains this ARE. In other words, we show that the lower bound of $\pi/2$ on the ARE is tight, and that it is attained regardless of the particular realization of $\theta$ or the radius of the parameter space from which it is taken. Clearly, the minimal penalty on the efficiency of $\pi/2$ also holds under setting (iii), although the question whether this efficiency is achievable (or otherwise, what is the minimal ARE) remains open. 
\\

@@

We also note that our settings (ii) and (iii) are special cases of 
\cite{zhang2013information} that consider adaptive and distributed estimation protocols for $m$ machines, each has access to $n/m$ independent samples. The main result of \cite{zhang2013information} are bounds on the estimation error as a function of the number of bits $R$ each machine uses for communication. The specialization of their result to our setting, by taking $m=n$ and $R=1$, leads to looser lower bounds then $\sigma^2\pi/(2n)$ for cases (ii) and (iii). The counterpart of our setting (iii) in the case of hypothesis testing was considered in \cite{52470}, although the results there cannot be extended to parametric estimation. Other related works include statistical inference under multiterminal data compression \cite{han1987hypothesis, zhang1988estimation}, and one-bit quantization constraints in compressed sensing  \cite{baraniuk2017exponential} and in MIMO detection in wireless communication \cite{singh2009limits}. \\

@@

In this work we are concerned with the Bayes MSE risk defined as
\begin{equation}
\label{eq:error_def}
\mathbb E\left(\widehat{\theta}_n - \theta \right)^2,
\end{equation}
where the expectation is taken with respect to the distribution of $X^n$ and the prior distribution $\pi(\theta)$. \\

The main problem we consider is the minimization of \eqref{eq:error_def} over all encoding and estimation strategies and the characterization of its minimal value as a function of $n$. This minimization is the combination of the following two procedures: (1) selecting the $i$th message $M_i$ based on past messages and current observation $X_i$, and (2) estimating $\theta$ given messages $M^n$. 
%
We are interested in particular on the increase in sample complexity compared to the vanilla mean estimation without one-bit constraint. For this reason, we consider 
\begin{equation}
{\sigma^2} n \mathbb E\left(\widehat{\theta}_n - \theta \right)^2
\label{eq:relative_efficiency}
\end{equation}
in the limit as $n$ goes to infinity. Equation \eqref{eq:relative_efficiency} is the ratio between the MSE attained by the empirical mean of the samples and the MSE attained by the estimator $\widehat{\theta}_n$. Note that if the limit $\mathbb E\left(\widehat{\theta}_{n} - \theta \right)^2$ exists and finite, than \eqref{eq:relative_efficiency} is the ARE of $\widehat{\theta}_n$ \cite[Def. 6.6]{lehmann2006theory}.

@@

Before deriving our main results, we comment on the relation between our setting and the remote multiterminal source coding problem, also known as the CEO problem.

\section{Relation to Remote Multiterminal Source Coding \label{sec:ceo}}
%In this section we draw a connection between our general problem of mean estimation from one-bit samples and the remote multiterminal source coding problem \cite{berger1996ceo}, also known as the CEO problem. 
The setting of the CEO includes $n$ encoders, each has access to a noisy version of a random source sequence \cite{berger1996ceo}. 
% $\left\{ U_j,\,j=1,2,\ldots \right\}$. 
The $i$th encoder observes $k$ noisy source symbols and transmit $R_i k$ bits to a central estimator. \par
Assuming that $\theta$ is drawn once from the prior $\pi(\theta)$, our mean estimation problem from one-bit samples under distributed encoding (setting (iii) in the Introduction) corresponds to the Gaussian CEO setting with $k=1$ source realization: the $i$th encoder uses $R_i=1$ bits to transmit a message that is a function of $X_i = \theta + \sigma Z_i$, where $Z_i$ is standard normal. As a result, a lower bound on the MSE distortion in estimating $\theta$ in the distributed encoding setting is given by the MSE in the optimal source coding scheme for the CEO with: $n$ terminals of codes rates $R_1 = \ldots = R_n = 1$, a Gaussian observation noise at each terminal of variance $\sigma^2$, and an arbitrary number of $k$ independent draws of $\theta$. Note that the difference between the CEO and ours lays in the privilege of each of the CEO encoders to describe $k$ realizations of $\theta$ using $k$ bits with MSE averaged over these realization, whereas our setting only allows $k=1$. 
 \\

@@

\begin{proof}
We consider the expression \cite[Eq. 10]{chen2004upper} that gives the minimal distortion $D^\star$ in the CEO with $L$ observers and under a total sum-rate $R_\Sigma = R_1 + \ldots +R_L$:
\begin{equation} \label{eq:ceo_optimal_sumrate}
R_{\Sigma} = \frac{1}{2} \log^+ \left[ \frac{\sigma_\theta^2}{D^\star} \left( \frac{D^\star L}{ D^\star L - \sigma^2 + D^\star \sigma^2 / \sigma_\theta^2 }\right)^L  \right].
\end{equation}
Assuming $R_\Sigma = n$ and $L=n$, we get
\begin{equation} \label{eq:ceo_optimal_sumrate}
n = \frac{1}{2} \log_2 \left[ \frac{\sigma_\theta^2}{D^\star} \left(\frac{ D^\star n }{D^\star n - \sigma^2 + D^\star \sigma^2/\sigma_\theta^2 }  \right)^n  \right].
\end{equation}
The value of $D^\star$ that satisfies the equation above describes the MSE under an optimal allocation of the sum-rate $R_\Sigma = n$ among the $n$ encoders. Therefore, $D^\star$ provides a lower bound to the CEO distortion with $R_1=\ldots,R_n = 1$ and hence a lower bound to the minimal MSE in estimating $\theta$ in the distributed encoding setting. By considering $D^\star$ in \eqref{eq:ceo_optimal_sumrate} as $n\rightarrow \infty$, we see that 
\[
D^\star = \frac{ 4\sigma^2 }{3n + 4 \sigma^2 / \sigma_\theta^2 } + o(n^{-1}) =  \frac{4\sigma^2}{3n} + o(n^{-1}). 
\]
\end{proof}

Prop.~\ref{prop:ceo_lower_bound} implies that, unlike in the centralized setting, there is a loss in efficiency in estimating $\theta$ due to one-bit measurements in this setting. In the next section we show that the ARE in adaptive encoding setting does not exceeds $\pi/2$, and thus provides a tighter lower bound for the distributed encoding setting than $4/3$ of  \eqref{eq:ceo_bound}. \\

We note although the lower bound \eqref{eq:ceo_bound} was derived assuming the optimal allocation of $n$ bits per observation among the encoders, this bound cannot be tightened by considering the CEO distortion while enforcing the condition $R_1=\ldots = R_n = 1$. Indeed, an upper bound for the CEO distortion under the condition $R_1=\ldots = R_n = 1$ follows from \cite{KipnisRini2017}, and leads to
\[
D_{CEO} \leq  \left( \frac{1}{\sigma_\theta^2} +  \frac{3n}{4\sigma^2 + \sigma_\theta^2} \right)^{-1}   =
\frac{4 \sigma^2}{3n} +  \frac{\sigma_\theta^2}{3n} + O(n^{-2}),
\]
which is equivalent to \eqref{eq:ceo_bound} when $\sigma_\theta$ goes to zero. 
%Recall that the difference between the MSE lower bound \eqref{eq:ceo_bound} and the actual MSE in the distributed encoding setting is attributed exclusively to the ability to perform coding over blocks. Namely, the freedom to consider an arbitrary number $k$ of independent realizations of the problem, versus only one in ours. Therefore, since 
%In other words, it is the ability to exploit the geometry of a high-dimensional product space, rather than the distributed nature of the problem, that distinguishes between the CEO distortion and our problem of mean estimation from one-bit samples.\par

\section{Results \label{sec:sequential}}
The first main result of this paper, as described in Thm.~\ref{thm:adpative_lower_bound} below, states that the ARE of any adaptive estimator cannot be lower than $\pi/2$. Next, we provide a particular adaptive estimation scheme and show in Thm.~\ref{thm:sgd} that its efficiency is $\pi/2$. Finally, in Thm.~\ref{thm:opt_one_step}, we provide an adaptive estimation scheme that is one-step optimal in the sense that at each step $i$, the encoder send the message $M_i^\star$ that minimizes the MSE given $X_i$ and the previous $M^{i-1}$ messages. While it is not clear whether the efficiency of this last scheme is $\pi/2$, numerical simulations suggests that the MSE of this scheme times $n$ also converges to $\pi/2$ faster than the first scheme.

@@

\subsection{A lower bound on adaptive one-bit schemes}
Our first results asserts that the ARE \eqref{eq:relative_efficiency} of any adaptive estimation scheme is bounded from below by $\pi/2$, as follows from the following theorem:
\begin{thm}[minimal relative effeciency] \label{thm:adpative_lower_bound}
Let $\widehat{\theta}_n$ be any estimator of $\theta$ in the adaptive setting of Fig.~\ref{fig:sequential}. Assume that the density of the prior $\pi(\theta)$ converges to zero at the endpoints of the interval $\Theta$. Then
\[
\mathbb E\left[ (\theta-\theta_n)^2 \right] \geq  \frac{\pi \sigma^2 }{2n +\pi \sigma^2  I_0}  =  \frac{\pi}{2n}\sigma^2+O(n^{-2}),
\]
where 
\[
I_0 = \mathbb E \left( \frac{d}{d\theta} \log \pi (\theta) \right)^2
\]
is the Fisher information with respect to a location model in $\theta$. 
\end{thm}

@@

For the estimator defined by \eqref{eq:sgd_alg} and \eqref{eq:sgd_est} we have the following results:
\begin{thm} \label{thm:sgd}
The sequence $\widehat{\theta}_n$ of \eqref{eq:sgd_est} satisfies
\begin{enumerate}
\item[(i)]
\[
\sqrt{n} \left( \widehat{\theta}_n - \theta \right) \overset{d}{\rightarrow} \mathcal N \left(0,  \pi \sigma^2 /2 \right).
\]
\item[(ii)] In addition to the conditions above, assume that $\gamma_n = o(n^{-2/3})$ and $\sum_{n=1}^\infty \gamma_n = \infty$ (e.g., $\gamma_n = n^{-\beta}$ with $2/3<\beta<1$). Then
\[
\lim_{n\rightarrow \infty} n\mathbb E \left[ \left(\theta-\widehat{\theta}_n \right)^2 \right] = \frac{\pi}{2} \sigma^2 . 
\]
\end{enumerate}

@@

Thm.~\ref{thm:sgd} implies that the estimator $\widehat{\theta}_n$, defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg}, attains the minimal ARE as established by Thm.~\ref{thm:adpative_lower_bound}.
\par
Note that $\theta_0$ is not explicitly defined 
in equation \eqref{eq:sgd_est}. While a reasonable initialization is $\theta_0 = \mathbb E [\theta]$, Thm.~\ref{thm:sgd} implies that the asymptotic behavior of the estimator is indifferent to this initialization. Thus, the optimal efficiency is attained regardless of the prior distribution on $\theta$ or the radius of the parameter space $\Theta$. Nevertheless, the bound in Thm.~\ref{thm:adpative_lower_bound} suggests that the non-asymptotic estimation error can be significantly reduced whenever the location information $I_0$ is large. In contrast, the one-step optimal scheme presented in the following subsection updates the prior distribution on $\theta$ given all information gathered until step $n-1$ to provide the step $n$ estimate and prior. In particular, the this scheme exploit the prior information on $\theta$ provided by $\pi(\theta)$.

\subsection{One-step optimal estimation}
We now consider an estimation scheme that posses the property of \emph{one-step optimality}: at each step $i$, the $i$th encoder designs the detection region $M_i^{-1}(1)$ such that the MSE given $M^i$ is minimal. In other word, this scheme designs the messages $M^n$ in a greedy manner, such that the MSE at step $i$ is minimal given the current state of the estimation described by $M^{i-1}$. \\

@@

\begin{proof}
The proof is completed by the following two lemmas, proofs of which can be found in the Appendix:
\begin{lem} \label{lem:unique}
Let $f(x)$ be a log-concave probability density function. Then the equation 
\begin{equation}
\label{eq:lem_fixed_point}
2x = \frac{\int_x^\infty uf(u)du}{\int_x^\infty f(u)du} + \frac{\int_{-\infty}^x uf(u)du}{\int_{-\infty}^x f(u)du} 
\end{equation}
has a unique solution.
\end{lem}
\begin{lem} \label{lem:adaptive}
Let $U$ be a random variable with probability density function  $P(du)$. Then the one-bit message $M^\star\in \{-1,1\}$ that minimizes
\[
\int \left( u - \mathbb E[U|M(u)]  \right)^2 P(du)
\]
is given by
\[
M^\star  =  \sgn(U - \tau),
\]
where $\tau$ is the unique solution to
 \[
2 \tau = \frac{\int_{\tau}^\infty u P(du)} {\int_{\tau}^\infty P(du)} + \frac{\int_{-\infty}^{\tau} u P(du)}{\int_{-\infty}^{\tau} P(du)}.
\]
\end{lem}
\end{proof}

Thm.~\ref{thm:opt_one_step} suggests the following adaptive encoding and estimation scheme: 
\begin{itemize}
\item Initialization: set $P_0(t) = \pi(\theta)$.
\item For $n\geq 1$:
\begin{enumerate}
\item Update the prior as
\begin{align}
P_n(t) = & P(\theta=t |M^n) \\
& = \frac{ P\left( \theta=t | M^{n-1} \right) P(M_n | \theta = t , M^{n-1})  } { P(M_n | M^{n-1} )} \nonumber \\ 
& = \alpha_n  P_{n-1}(t) \Phi\left(M_n \frac{ t - \tau_{n-1} }{\sigma} \right), \label{eq:density_update}
\end{align}
where $\alpha_n$ is a normalization coefficient that equals
\[
\alpha_n = \left(\int_{\mathbb R} P_{n-1}(t) \Phi\left(M_n \frac{t- \tau_{n-1} }{\sigma} \right)  dt \right)^{-1}. 
\]
\item The $n$th estimate for $\theta$ is the conditional expectation of $\theta$ given $M^n$, namely
\begin{equation}
\theta_n = \mathbb E \left[ \theta| M^n\right] = \int_{-\infty}^\infty t P_n(t) dt. \label{eq:estimator_update}
\end{equation}
\item Solve equation \eqref{eq:fixed_point} with the updated prior $P_n(t)$ instead of $\pi(d\theta)$. Note that since the standard normal cdf $\Phi(x)$ is log-concave, the updated prior $P_n(t)$ remains log-concave and thus a unique solution to \eqref{eq:fixed_point} is guaranteed by Lem.~\ref{lem:unique}. 
\item Update the $(n+1)$th message as
\begin{equation}\label{eq:message_update}
M_{n+1} = \sgn(X_{n+1}-\tau_n)
\end{equation}
\end{enumerate}
\end{itemize}
Since equation \eqref{eq:fixed_point} has no analytic solution, it is hard to derive the asymptotic behavior of the estimator defined by \eqref{eq:estimator_update} and \eqref{eq:message_update}. We conjecture that it attains the asymptotic relative efficiency of $\sigma^2\pi/2$ as can be observed from the numerical simulation illustrated in Fig.~\ref{fig:adaptive_error}. Also shown in Fig.~\ref{fig:adaptive_error} are the normalized MSE of the asymptotically optimal estimator defined by \eqref{eq:sgd_alg} and \eqref{eq:sgd_est}, as well as the MSE achieved by the empirical mean of the samples for the same sample realization.

@@

\subsection*{Proof of Thm.~\ref{thm:opt_one_step}}
In this subsection we prove Lem. \ref{lem:adaptive} and \ref{lem:unique} that lead to Thm.~\ref{thm:opt_one_step}.
